---
title: "DATA 607 - Assignment 2 - SQL and R"
author: "Preston Peck"
date: "9/2/2021"
output: html_document
---

# Overview

A small group of participants of varying ages, occupational backgrounds, and viewing experiences was tasked with providing their opinions of films of varying ages, genres, and socially accepted numerical ratings. The following is a brief analysis of these results:



# Analysis
### Import Dependencies

```{r}
require("RPostgreSQL")
require("ggplot2")
```

### Database Connection
##### Password Protection (left intentionally visible)

```{r}
secret <- {
  "password"
}
```

##### Connect

```{r}
driver <- dbDriver("PostgreSQL")
connection <- dbConnect(driver, dbname = "movie",
                 host = "localhost", port = 5432,
                 user = "postgres", password = secret)
rm(secret)
```

### Database Query
##### What movies received the best reviews?

```{r}
averageStarsByMovieQuery <-
"select m.title, avg(r.num_of_star) as average_critic_star_rating 
from critic_movie as cm
left join critic as c
on c.id = cm.critic_id
left join rating as r
on r.id = cm.rating_id
left join movie as m
on m.id = cm.movie_id
group by m.id
order by average_critic_star_rating;"

averageStarsByMovie <- dbGetQuery(connection, averageStarsByMovieQuery)
averageStarsByMovie

averageAcrossAllMovies <- mean(averageStarsByMovie$average_critic_star_rating)
averageAcrossAllMovies
  
ggplot(averageStarsByMovie, aes(x = reorder(title, average_critic_star_rating), y = average_critic_star_rating, fill = average_critic_star_rating > averageAcrossAllMovies)) + 
  geom_col() + labs(x = "Movie", y = "Average Stars", fill = "Rating > Avg?") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##### What movie had the biggest critic to Metacritic difference?

```{r}
criticMetacriticStarRatingDifferenceByMovieQuery <-
"select title, average_critic_star_rating, num_of_star as metacritic_star_rating, (average_critic_star_rating - num_of_star) as critic_metacritic_star_rating_difference
from(
  select m.title, m.metacritic_rating_id, avg(r.num_of_star) as average_critic_star_rating 
  from critic_movie as cm
  left join critic as c
  on c.id = cm.critic_id
  left join rating as r
  on r.id = cm.rating_id
  left join movie as m
  on m.id = cm.movie_id
  group by m.id
  order by average_critic_star_rating
) as cmr
left join rating as r
on r.id = cmr.metacritic_rating_id;"

criticMetacriticStarRatingDifferenceByMovie <- dbGetQuery(connection, criticMetacriticStarRatingDifferenceByMovieQuery)
criticMetacriticStarRatingDifferenceByMovie

ggplot(criticMetacriticStarRatingDifferenceByMovie, aes(x = reorder(title, critic_metacritic_star_rating_difference), y = critic_metacritic_star_rating_difference, fill = critic_metacritic_star_rating_difference > 0)) +
  geom_col() + 
  labs(x = "Movie", y = "Average Stars", fill = "Critic > Metacritic?") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##### Which level of hobbyist gave the best reviews?

```{r}
averageStarsByExperienceQuery <- 
"select e.name, avg(r.num_of_star) as average_critic_star_rating 
from critic_movie as cm
left join critic as c
on c.id = cm.critic_id
left join rating as r
on r.id = cm.rating_id
left join experience as e
on e.id = c.experience_id
group by e.id
order by average_critic_star_rating;"

averageCriticScore <- dbGetQuery(connection, averageStarsByExperienceQuery)
averageCriticScore

ggplot(averageCriticScore, aes(x = reorder(name, average_critic_star_rating), y = average_critic_star_rating, fill = reorder(name, average_critic_star_rating))) +
  geom_col() + 
  scale_fill_manual(values = c("#FFFF00", "#FFA500", "#FF0000")) + 
  labs(x = "Viewing Experience", y = "Average Stars", fill = "Experience")
```

##### Which generation reviewed movies within their generation the best?

```{r}
averageStarsByCriticMovieGenerationMatchQuery <-
"select concat(c.generation, '_', m.generation) as client_movie_generation, avg(r.num_of_star) as average_critic_star_rating
from critic_movie as cm
left join(
  select *, case 
  when year_born <= 1945 then 'Silent Gen'
  when year_born >= 1946 and year_born <= 1964 then 'Baby Boomers'
  when year_born >= 1965 and year_born <= 1976 then 'Gen X'
  when year_born >= 1977 and year_born <= 1995 then 'Millennials'
  when year_born >= 1996 and year_born <= 2015 then 'Gen Z'
  else 'Gen Alpha'
  end generation
  from(
  	select *, extract(year from birthday) as year_born 
  	from critic
  ) critic_year_born
) as c
on c.id = cm.critic_id
left join rating as r
on r.id = cm.rating_id
left join(
  select *, case 
  when year_released <= 1945 then 'Silent Gen'
  when year_released >= 1946 and year_released <= 1964 then 'Baby Boomers'
  when year_released >= 1965 and year_released <= 1976 then 'Gen X'
  when year_released >= 1977 and year_released <= 1995 then 'Millennials'
  when year_released >= 1996 and year_released <= 2015 then 'Gen Z'
  else 'Gen Alpha'
  end generation
  from movie
) as m
on m.id = cm.movie_id
where c.generation = m.generation
group by c.generation, m.generation
order by average_critic_star_rating;"

averageStarsByCriticMovieGenerationMatch <- dbGetQuery(connection, averageStarsByCriticMovieGenerationMatchQuery)
averageStarsByCriticMovieGenerationMatch

ggplot(averageStarsByCriticMovieGenerationMatch, aes(x = reorder(client_movie_generation, average_critic_star_rating), y = average_critic_star_rating)) +
  geom_col() +
  labs(x = "Critic to Movie Generation", y = "Average Stars")
```

##### How did generations review movies outside their generation?
```{r}
averageStarsByCriticMovieGenerationMismatchQuery <-
"select concat(c.generation, '_', m.generation) as client_movie_generation, avg(r.num_of_star) as average_critic_star_rating
from critic_movie as cm
left join(
  select *, case 
  when year_born <= 1945 then 'Silent Gen'
  when year_born >= 1946 and year_born <= 1964 then 'Baby Boomers'
  when year_born >= 1965 and year_born <= 1976 then 'Gen X'
  when year_born >= 1977 and year_born <= 1995 then 'Millennials'
  when year_born >= 1996 and year_born <= 2015 then 'Gen Z'
  else 'Gen Alpha'
  end generation
  from (
  	select *, extract(year from birthday) as year_born 
  	from critic
  ) critic_year_born
) as c
on c.id = cm.critic_id
left join rating as r
on r.id = cm.rating_id
left join(
  select *, case 
  when year_released <= 1945 then 'Silent Gen'
  when year_released >= 1946 and year_released <= 1964 then 'Baby Boomers'
  when year_released >= 1965 and year_released <= 1976 then 'Gen X'
  when year_released >= 1977 and year_released <= 1995 then 'Millennials'
  when year_released >= 1996 and year_released <= 2015 then 'Gen Z'
  else 'Gen Alpha'
  end generation
  from movie
) as m
on m.id = cm.movie_id
where c.generation != m.generation
group by c.generation, m.generation
order by average_critic_star_rating;"

averageStarsByCriticMovieGenerationMismatch <- dbGetQuery(connection, averageStarsByCriticMovieGenerationMismatchQuery)
averageStarsByCriticMovieGenerationMismatch

ggplot(averageStarsByCriticMovieGenerationMismatch, aes(x = reorder(client_movie_generation, average_critic_star_rating), y = average_critic_star_rating)) +
  geom_col() +
  labs(x = "Critic to Movie Generation", y = "Average Stars") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



# Conclusions

It's hard to make any strong conclusions here given the limited sample size, but on the face, the best one that can be made is that movie enthusiasts seem more likely to rate movies higher on average than those who are not as familiar with the art.